Final Project

### 2025-05-02 MSSC5750_Project.qmd
Fixing Shap Values

Shap values/viz
```{r}
# Requires split out xtrain, ytrain, xtest... Also a combined x/ytrain

# We will apply to lasso data

# We need to manually generate the dummy variables and match naems to what the model created
# coder<-dummyVars(~ person_home_ownership + loan_intent + previous_loan_defaults_on_file, data = data, fullRank = TRUE, sep='')
# trainDataAdj<-cbind(trainData %>% select(-person_home_ownership, -loan_intent, -previous_loan_defaults_on_file), predict(coder, newdata = trainData) )
# testDataAdj<-cbind(testData %>% select(-person_home_ownership, -loan_intent, -previous_loan_defaults_on_file), predict(coder, newdata = testData) )

# Only want the columns that we included in lasso

xtrain<-trainData %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file)
ytrain<-trainData %>% select(loan_status)
xytrain<-trainData %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file,loan_status)
xtest<-testData  %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file)

@@

# we will just explain on our train data as well for larger sample size
set.seed(5)
xtrainSmall <- xtrain[sample(nrow(xtrain), 1000), ]
explainer<- shapr::explain(model=logModelLasso, x_train=xtrainSmall, x_explain=xtrainSmall, approach="ctree", phi0=mean(ytrain$loan_status))

@@

###
shapObj <-shapviz(explainer)
sv_importance(shapObj, kind="beeswarm", show_numbers = TRUE, max_display=15)

>>>>

Shap values/viz
```{r}
# Requires split out xtrain, ytrain, xtest.

xtrain<-trainDataBalanced %>% select(-c("lassoPred", "person_income", "person_emp_exp", "previous_loan_defaults_on_file_Yes", "loan_status"))
ytrain<-trainDataBalanced %>% select(loan_status)

xtest<-testData  %>% select(-c("lassoPred", "person_income", "person_emp_exp", "previous_loan_defaults_on_file_Yes", "loan_status", "logisticPred","logisticLassoPred", "rebalancedLog"))

@@

# we will just explain on our train data as well for larger sample size
set.seed(5)
xtestSmall <- xtest[sample(nrow(xtest), 250), ]
explainer<- shapr::explain(model=rebalancedLog, x_train=xtrain, x_explain=xtestSmall, approach="ctree", phi0=mean(ytrain$loan_status))

@@

###
shapObj <-shapviz(explainer)
sv_importance(shapObj, kind="beeswarm", show_numbers = TRUE, max_display=25)
sv_importance(shapObj, kind="beeswarm", show_numbers = TRUE, max_display=10)
